Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
update text in the power cmp

For non-linearity patterns, the power curve of RESET test climbs aggressively from
`r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(0), offset0 = log(0.05/0.95)), type = "response")) * 100)`% to `r round(unname(predict(reset_poly_glmmod, newdata = data.frame(effect_size = exp(2.5), offset0 = log(0.05/0.95)), type = "response")) * 100)`% as $log_e(E)$ increases from 0 to 2.5, while power of other tests respond inactively to the change of effect, showing that RESET test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable residual departures can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test is almost always greater than the power of visual test. For $log_e(E)$ at around 2.5, where the power curve of the visual test remains at a low level, the Breusch-Pagan test has around 50% chance of rejecting $H_0$. Similarly, the visual feature is nearly unobservable from the example residual plots.

The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features. This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The visual pattern becomes much clearer as $log_e(E)$ increases. At $log_e(E) = 5$, the power reaches almost 100%.

The power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

@@

target_b <- c(0.54, 1.48, 4.02, 10.92, 29.68, 80.69, 219.33)

# Calculate the effect size 
es <- map_dbl(target_b, 
              function(.x){
                mod <- heter_model(a = 1, 
                                   b = .x, 
                                   x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})
                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
                }
              )
```

@@

# 10091
# 101
# 105
# 106
# 112
# 119
# 133
set.seed(135)

@@

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 1, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

>>>>

For both the non-linearity pattern and the heteroskedasticity pattern, the power curves of appropriate conventional tests are constantly greater than the power curves of visual tests and other inappropriate conventional tests. At $log(E) = 1$ for non-linearity model, where the residual departures are unnoticeable from the corresponding residual plot, the RESET test has above 25% of chance of rejecting the null hypothesis, while only a few visual tests would result in a rejection. Similarly, at $log(E) = 3$ for heteroskedasticity model, the BP test has around 25% power but the non-constant variance visual pattern is almost unobserved from the residual plot and the power of the visual test is close to the significance level at this point.

Further, the power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

@@

target_b <- c(0.27, 0.74, 2.01, 5.46, 14.84, 40.34, 109.66)

# Calculate the effect size 
es <- map_dbl(target_b, 
              function(.x){
                mod <- heter_model(a = 0, 
                                   b = .x, 
                                   x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})
                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
                }
              )
```

@@

# 10091
# 101
# 105
# 106
# 112
# 119
# 133
set.seed(100)

@@

# Obtain data at different b
for (i in target_b) {
  j <- j + 1
  plist[[j]] <- heter_model(a = 0, 
                            b = i, 
                            x = rand_uniform(-1, 1))$
                  gen(n = 100, computed = select(ex_dat, x, e)) %>%
  VI_MODEL$plot(theme = theme_light(base_size =  5),
                remove_axis = TRUE,
                remove_grid_line = TRUE, size = 0.25, stroke = 0.25) +
    ggtitle(round(es[j], 1)) +
    theme(aspect.ratio = 0.8)
}

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
update text in the power cmp

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 10\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
while power of other tests respond inactively to the change of effect,
showing that RESET test is way more sensitive to the type of model
defects that being considered. Meanwhile, no noticeable residual
departures can be spotted from the example residual plots.

In terms of heteroskedasticity patterns, the power of Breusch-Pagan test
is almost always greater than the power of visual test. For \(log_e(E)\)
at around 2.5, where the power curve of the visual test remains at a low
level, the Breusch-Pagan test has around 50\% chance of rejecting
\(H_0\). Similarly, the visual feature is nearly unobservable from the
example residual plots.

The power of visual test arises steadily as \(log_e(E)\) increases from
3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
suggesting that the effect starts to make significant impact on the
degree of the presence of the designed visual features. This can also be
observed from the example residual plots that when \(log_e(E) = 3.5\), a
weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The
visual pattern becomes much clearer as \(log_e(E)\) increases. At
\(log_e(E) = 5\), the power reaches almost 100\%.

The power of all inappropriate tests show improvement as the effect
increases but at a lower rate than the visual test in both scenarios.
This coincides the point made by \citet{cook1982residuals} that
residual-based tests for a specific type of model defect may be
sensitive to other types of model defects. The power curve of RESET test
remains at around 5\% in Figure \ref{fig:heterpower} since there are no
non-linear terms leave out in the heteroskedasticity model and \(H_0\)
of the test is always satisfied.

>>>>

For both the non-linearity pattern and the heteroskedasticity pattern,
the power curves of appropriate conventional tests are constantly
greater than the power curves of visual tests and other inappropriate
conventional tests. At \(log(E) = 1\) for non-linearity model, where the
residual departures are unnoticeable from the corresponding residual
plot, the RESET test has above 25\% of chance of rejecting the null
hypothesis, while only a few visual tests would result in a rejection.
Similarly, at \(log(E) = 3\) for heteroskedasticity model, the BP test
has around 25\% power but the non-constant variance visual pattern is
almost unobserved from the residual plot and the power of the visual
test is close to the significance level at this point.

Further, the power of all inappropriate tests show improvement as the
effect increases but at a lower rate than the visual test in both
scenarios. This coincides the point made by \citet{cook1982residuals}
that residual-based tests for a specific type of model defect may be
sensitive to other types of model defects. The power curve of RESET test
remains at around 5\% in Figure \ref{fig:heterpower} since there are no
non-linear terms leave out in the heteroskedasticity model and \(H_0\)
of the test is always satisfied.
